Conversations in time: interactive visualisation to explore structured temporal data

### 2020-09-26 tsibbletalk.Rmd
minor tweaks

<!-- In Figure \ref{fig:tourism-linking-fig}, when hovering and selecting the circle associated with "Sydney" in the scatter plot, all data records with shared values of "Sydney" listen and react to this interaction via self updating in red. In order for comparison with other regions or states, press the "Shift" key to enable persistent selection, and simultaneously select the parent node on the tree, saying "Western Australia", to include all the children by switching to the blue colour.-->
The domestic tourism sees Sydney as one of the most popular destinations in realm of business and friends visiting over the years. Despite the relatively weaker performance in Western Australia, Australia's North West region sees a strongest upward trend in Business, bypassing Sydney in some years.

@@

An important aspect of temporal data is the time context. Time has cyclical structure, that may correspond to seasonal patterns to be discovered. The `index` component of the shared tsibble data forms the basis for exploring seasonality. To investigate for periodic or aperiodic patterns, series should be wrapped on themselves, where the index is broken into temporal components like quarter or day. We shall explore this with pedestrian traffic in Melbourne, Australia.

<!-- The shared tsibble data leverages the `key` attribute to converse with many coordinated views, with or without \CRANpkg{shiny}. On the other hand, a second critical attribute--`index`--lays the foundational temporal context that augments the conversation. When temporal data are plotted and stretched against the entire span like Figure \ref{fig:highlight-retail-1}, it puts emphasis on the trend perception. Yet to digest periodic/aperiodic patterns, data should be wrapped over relative time units that are origin-less, such as one quarter or one day.-->

```{r wrap-ped, fig.show = "hold", fig.cap = "Snapshots wrapping after slicing the \\code{pedestrian20} data at different intervals, none (top), daily (middle) and weekly (bottom). This type of interaction is made possible with Shiny elements.", fig.subcap = c("Initial overview state", "1-day state", "7-day state, anchoring to Monday"), fig.ncol = 1}
include_graphics("img/wrap-0.png")
include_graphics("img/wrap-1.png")
include_graphics("img/wrap-7.png")
```

The city of Melbourne has sensors installed at various locations, to record hourly counts of pedestrians, in order to capture daily rhythms of the downtown [@ped]. Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at four different locations, for three different time slices, daily, weekly and full five months. Plot \ref{fig:wrap-ped-1} shows hourly counts from January to May on an absolute time line, facetted by location. The stage 3 COVID-19 lockdown, on March 16, is marked by a change of colour. (The pre-lockdown period is coloured with dark green and lockdown with orange.) We can see a significant decline in foot traffic  at all four locations. QV Market is less affected probably because this is a major produce market, an essential service that continued to operate. Bourke St, a major shopping center, sees a gradual uptick in the last weeks of the period indicating that people were getting back into the shops.

>>>>

The domestic tourism sees Sydney as one of the most popular destinations in realm of business and friends visiting over the years. Despite the relatively weaker performance in Western Australia, Australia's North West region sees a strongest upward trend in Business, bypassing Sydney in some years.

@@

An important aspect of temporal data is the time context. Time has cyclical structure, that may correspond to seasonal patterns to be discovered. The `index` component of the (shared) tsibble data forms the basis for exploring seasonality. To investigate for periodic or aperiodic patterns, series should be wrapped on themselves, where the index is broken into temporal components like quarter or day. We shall explore this with pedestrian traffic in Melbourne, Australia.

```{r wrap-ped, fig.show = "hold", fig.cap = "Snapshots wrapping after slicing the \\code{pedestrian20} data at different intervals, (a) none, (b) daily and (c) weekly. This type of interaction is made possible with Shiny elements.", fig.subcap = c("Initial overview state", "1-day state", "7-day state, anchoring to Monday"), fig.ncol = 1}
include_graphics("img/wrap-0.png")
include_graphics("img/wrap-1.png")
include_graphics("img/wrap-7.png")
```

The city of Melbourne has sensors installed at various locations, to record hourly counts of pedestrians, in order to capture daily rhythms of the downtown [@ped]. Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at four different locations, for three different time slices, daily, weekly and full five months. Plot \ref{fig:wrap-ped-1} shows hourly counts from January to May on an absolute time line, facetted by locations. The stage 3 COVID-19 lockdown, on March 16, is marked by a change of colour. (The pre-lockdown period is coloured with dark green and lockdown with orange.) We can see a significant decline in foot traffic  at all four locations. QV Market is less affected probably because this is a major produce market, an essential service that continued to operate. Bourke St, a major shopping center, sees a gradual uptick in the last weeks of the period indicating that people were getting back into the shops.

### 2020-09-26 tsibbletalk.tex
minor tweaks

An important aspect of temporal data is the time context. Time has
cyclical structure, that may correspond to seasonal patterns to be
discovered. The \texttt{index} component of the shared tsibble data
forms the basis for exploring seasonality. To investigate for periodic
or aperiodic patterns, series should be wrapped on themselves, where the
index is broken into temporal components like quarter or day. We shall
explore this with pedestrian traffic in Melbourne, Australia.

@@

\caption[Snapshots wrapping after slicing the \code{pedestrian20} data at different intervals, none (top), daily (middle) and weekly (bottom)]{Snapshots wrapping after slicing the \code{pedestrian20} data at different intervals, none (top), daily (middle) and weekly (bottom). This type of interaction is made possible with Shiny elements.}\label{fig:wrap-ped}
\end{figure}
\end{Schunk}

The city of Melbourne has sensors installed at various locations, to
record hourly counts of pedestrians, in order to capture daily rhythms
of the downtown \citep{ped}. Figure \ref{fig:wrap-ped} shows the first
five months of 2020 foot traffic at four different locations, for three
different time slices, daily, weekly and full five months. Plot
\ref{fig:wrap-ped-1} shows hourly counts from January to May on an
absolute time line, facetted by location. The stage 3 COVID-19 lockdown,
on March 16, is marked by a change of colour. (The pre-lockdown period
is coloured with dark green and lockdown with orange.) We can see a
significant decline in foot traffic at all four locations. QV Market is
less affected probably because this is a major produce market, an
essential service that continued to operate. Bourke St, a major shopping
center, sees a gradual uptick in the last weeks of the period indicating
that people were getting back into the shops.

>>>>

An important aspect of temporal data is the time context. Time has
cyclical structure, that may correspond to seasonal patterns to be
discovered. The \texttt{index} component of the (shared) tsibble data
forms the basis for exploring seasonality. To investigate for periodic
or aperiodic patterns, series should be wrapped on themselves, where the
index is broken into temporal components like quarter or day. We shall
explore this with pedestrian traffic in Melbourne, Australia.

@@

\caption[Snapshots wrapping after slicing the \code{pedestrian20} data at different intervals, (a) none, (b) daily and (c) weekly]{Snapshots wrapping after slicing the \code{pedestrian20} data at different intervals, (a) none, (b) daily and (c) weekly. This type of interaction is made possible with Shiny elements.}\label{fig:wrap-ped}
\end{figure}
\end{Schunk}

The city of Melbourne has sensors installed at various locations, to
record hourly counts of pedestrians, in order to capture daily rhythms
of the downtown \citep{ped}. Figure \ref{fig:wrap-ped} shows the first
five months of 2020 foot traffic at four different locations, for three
different time slices, daily, weekly and full five months. Plot
\ref{fig:wrap-ped-1} shows hourly counts from January to May on an
absolute time line, facetted by locations. The stage 3 COVID-19
lockdown, on March 16, is marked by a change of colour. (The
pre-lockdown period is coloured with dark green and lockdown with
orange.) We can see a significant decline in foot traffic at all four
locations. QV Market is less affected probably because this is a major
produce market, an essential service that continued to operate. Bourke
St, a major shopping center, sees a gradual uptick in the last weeks of
the period indicating that people were getting back into the shops.
